---
title: "tutorial3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

## (a)

\begin{align*}
  F_X(x) &= \int_{-\infty}^x f_X(t) dt\\
  &= \int_{0}^x 2 e^{-2t} dt\\
  &= - e^{-2t} \,\Big\vert_0^x\\
  &= 1 - e^{-2x} \,.
\end{align*}

Notice that I use $t$ as the dummy variable inside the integral because $x$ is used as the upper limit in the definite integral.

## (b)

\begin{align*}
  \mathbb{E}(X) &= \int_{-\infty}^{\infty} xf_X(x) dx\\
  &= \int_{0}^{\infty}x f_X(x) dx\\
  &= \int_0^\infty \underbrace{x}_{u} \underbrace{e^{-2x} dx}_{dv}\qquad \text{integration by parts}\\
  &= -x e^{-2x} \,\Big\vert_0^\infty - \int_0^\infty - e^{-2x} dx\\
  &= - \frac{1}{2} e^{-2x} \,\Big\vert_0^\infty\\
  &= - \frac{1}{2} \, .
\end{align*}

Note: $-x e^{-2x} \,\Big\vert_0^\infty$ evaluates to 0, where l'Hopital's rule is used implicitly:

\begin{align*}
  & \, \,\lim_{c \rightarrow \infty} x e^{-2x}\\
  &= \lim_{c \rightarrow \infty} \frac{x}{e^{2x}} \text{ indeterminate form of a quotient of the form} \frac{\infty}{\infty}\\
  &= \lim_{c \rightarrow \infty} \frac{1}{2e^{2x}}\\
  &= 0 \,.
\end{align*}

## (c)

```{r}
set.seed(123)
u <- runif(10000, 0, 1)
x <- -0.5*log(1 - u)

hist(x, breaks=30)
```

To verify that our samples in the vector `x` are indeed drawn from $f_X(x)$, we can plot the latter distribution over the histogram. Note: notice the `prob=TRUE` argument in hist that renders turns the histogram into an approximate density.

```{r}
pdf <- function(z){
  return (2*exp(-2*z))
}

z <- seq(0, 3, by=0.001)

hist(x, breaks=30, prob=TRUE)
lines(z, pdf(z))
```

## (d)

```{r}
mean(x)
```

The above is close to 0.5, evaluated in (b). Looks good to me!

# Question 2

## (a)

Solve for $\int_0^2 f_Y(y) dy = 1$:
\begin{align*}
  &\,\,\int_0^2 f_Y(y) dy\\
  &= \int_0^2 c(2 - y) dy\\
  &= c\left(2y - \frac{y^2}{2} \right)\, \Bigg\vert^2_0\\
  &= 2c \, ,
\end{align*}
which implies that $c = \frac{1}{2}$.

## (b)

Using evidence from part **(a)** above, $\int f_Y(y) dy = y - \frac{y^2}{4}$. Now, we add the appropriate intervals for $y$ since $F_Y(y)$ needs to be written as a piecewise function.

\begin{align*}
  F_y(y) = \begin{cases}
    0 &\qquad \text{if } y < 0\\
    y - \frac{y^2}{4} &\qquad \text{if } 0 \leq y \leq 2\\
    1 &\qquad \text{if } y > 2 \, .
  \end{cases}
\end{align*}

## (c)

Notice that $P(1 < Y \leq 2) = P(Y \leq 2) - P(Y \leq 1) = 1 - P(Y \leq 1)$. Since we have the CDF from part **(b)**, we have that:
\begin{align*}
  P(1 < Y \leq 2) = 1 - P(Y \leq 1) = 1 - \left(1 - \frac{1}{4}\right) = \frac{1}{4} \, .
\end{align*}

## (d)

\begin{align*}
  \mathbb{E}(Y) &= \int_0^2 y \frac{1}{2}(2 - y) dy\\
  &= \frac{1}{2} \int_0^2 (2y - y^2) dy\\
  &= \frac{1}{2} \left(y^2 - \frac{y^3}{3}\right) \, \Bigg\vert^2_0\\
  &= \frac{2}{3}
\end{align*}

## (e)

\begin{align*}
  \mathbb{E}\left(Y^2\right) &= \int_0^2 y^2 \frac{1}{2}(2 - y) dy\\
  &= \frac{1}{2} \int_0^2 (2y^2 - y^3) dy\\
  &= \frac{1}{2} \left(\frac{2y^3}{3} - \frac{y^4}{4}\right) \, \Bigg\vert^2_0\\
  &= \frac{2}{3}
\end{align*}

We can find the variance by using the first and second moments:

\begin{align*}
  \operatorname{Var}(Y) = \frac{2}{3} - \left(\frac{2}{3}\right)^2 = \frac{2}{9}
\end{align*}